---
title: "Prelim Project Look - GroupF"
author: "Jocelyn, Daniel"
date: "3/1/2018"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

### Project Title: Travel in Developed and Developing Countries

```{r, include = FALSE}
library(tidyverse)
require(mosaic)
require(NSM3)
require(readr)
```

## Data Summary

### Data Read-In:
```{r}
data <- read.csv("Data/final_data.csv") %>% .[2:ncol(.)]
data %>% dim()
data %>% head()
```

### Summary Command:
```{r}
summary(data)
```

## Data Codebook

- airline
- airline.ID
- source.airport
- source.airport.id 
- destination.apirport
- destination.airport.id
- codeshare ("Y" if codeshare, empty otherwise)
- stops ("0" for direct flight)
- equipment (3-letter codes for plane types generally used on the flight, seperated by spaces if more than one)
- source.airport.name
- source.country     
- destination.airport.name
- destination.country
- Country.Code.x          
- X2010.x:X2016.x (GDP each year from 2010 - 2016)
- developed.x
- growth10_11.x:growth15_16.x (GDP growth between two years from 2010 - 2016)
- avg_growth.x
- Country.Code.y  
- X2010.y:X2016.y (GDP each year from 2010 - 2016)
- developed.y 
- growth10_11.y:growth15_16.y (GDP growth between two years from 2010 - 2016)
- avg_growth.y

Please note that a variable name with a '.x' corresponds to the source country of the flight and a variable name with a '.y' correpsonds to the destination country of the flight. Additionally, `developed` is a categorical variable that indicates whether a country is developing or developed using the GDP cutoff of 10,490 dollars. This value is inflation adjusted from $12,000 today's dollars.

### Selected Variable Analysis

#### Countries

```{r}
data$source.country %>% unique() %>% length()
data$destination.country %>% unique() %>% length()
```

156 unique source and destination countries.

```{r}
length(which(data$developed.x==1))/nrow(data)
```

Of which 68% of the outbound countries are developed.

```{r}
length(which(data$developed.y==1))/nrow(data)
```

And 68% of the inbound countries are developed.


### Distributions

#### Outbound Flights

```{r}
data %>% group_by(source.country) %>% summarize(count=n())  %>% 
  ggplot(aes(count)) + geom_histogram() + xlab("number of outbound flights for a country")
```

#### Inbound Flights
```{r}
data %>% group_by(destination.country) %>% summarize(count=n())  %>% 
  ggplot(aes(count)) + geom_histogram() + xlab("number of inbound flights for a country")
```

#### Developed vs Developing Countries Outbound Flights

```{r}
data %>% group_by(developed.x) %>% summarize(count=n()) %>% ggplot(aes(as.factor(developed.x),count)) + geom_bar(stat='identity') + xlab("developed (1) vs developing (0) countries") + ggtitle("outbound flights for developing vs developed countries")
```

## Analysis Plan


```{r}
data %>% group_by(source.country, avg_growth.x) %>% summarize(count=n()) -> temp
```

```{r}
temp %>% ggplot(aes(avg_growth.x, count)) + geom_point()
```

