---
title: "Analysis"
author: "Jocelyn, Daniel"
date: "3/1/2018"
output: pdf_document
---

```{r}
library(tidyverse)
```

## Data Summary

```{r}
data <- read.csv("data/final_data.csv") %>% .[2:ncol(data)]
data %>% dim()
data %>% head()
```

## Data Codebook

```{r}
data %>% names()
```

- airline
- airline.ID
- source.airport
- source.airport.id 
- destination.apirport
- destination.airport.id
- codeshare 
- stops 
- equipment
- source.airport.name
- source.country     
- destination.airport.name
- destination.country
- Country.Code.x          
- X2010.x:X2016.x
- developed.x
- Country.Code.y          
- X2010.y:X2016.y
- developed.y

### Selected Variable Analysis

#### Outbound Flights

```{r}
data %>% group_by(source.country) %>% summarize(count=n())  %>% 
  ggplot(aes(count)) + geom_histogram()
```

#### Inbound Flights
```{r}
data %>% group_by(destination.country) %>% summarize(count=n())  %>% 
  ggplot(aes(count)) + geom_histogram()
```

#### Developed vs Developing Countries Outbound Flights

```{r}
data %>% group_by(developed.x) %>% summarize(count=n()) %>% ggplot(aes(as.factor(developed.x),count)) + geom_bar(stat='identity')
```

## Analysis Plan