---
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
---
### STAT 225 - Data Read In - Group F

```{r, include=FALSE}
require(mosaic)
library(tidyverse)
```

Our project is about exploring differences in  travel, via  plane, between developed and developing  countries.

### Data Sources

Our data sources are: 
- the Flight Routes Database from Kaggle [link](kaggle.com/open-flights/flight-route-database),
- airport codes from [airportcodes.org](airportcodes.org), and
- national Per Capita GDP figures from the World Bank [link](data.worldbank.org/indicator/NY.GDP.PCAP.PP.CD).

### Workflow to Generate your Data Set

```{r}
data <- read.csv("Data/routes.csv")
codes <- read.csv("Data/codes.csv", col.names="data")
gdp_data <- read_csv("Data/gdp_data.csv") %>% data.frame()
codes %>% extract(col=data, c("Airport", "Airport_Code"), "(.+)\\((...)\\)") -> codes
sapply(codes$Airport, function(x){unlist(strsplit(x,'-',''))[[1]]}) -> codes$Airport
codes %>% extract(col=Airport, c("Airport", "Country"), "(.+), (.+) ") -> codes
na.omit(codes) -> codes; nrow(codes)
```

Lose some codes 2508->2477

```{r}
temp <- left_join(data, codes, by = c("source.airport" = "Airport_Code")) %>%
  filter(!is.na(Country))

colnames(temp)[10:11] <- c("source.airport.name","source.country")

final_data <- left_join(temp, codes, by = c("destination.apirport" = "Airport_Code")) %>% 
  filter(!is.na(Country))

colnames(final_data)[12:13] <- c("destination.airport.name","destination.country")

nrow(final_data)
```

67663 -> 42495

```{r}
sapply(gdp_data[,3:9], function(x){which(is.na(x)) %>% length()})
is.na(gdp_data$X2010) -> missing
gdp_data[!missing,] -> gdp_data; nrow(gdp_data)
```

We have the most data for GDP  (PPP) figures for year 2010, so we'll proceed with this year. We now have n=238.

Using the inflation calculator provided by the BLS, https://data.bls.gov/cgi-bin/cpicalc.pl. \$12000 today would have been worth \$10490.48 in 2010.

```{r}
gdp_data$developed <- ifelse(gdp_data$X2010>10490.48, 1, 0); gdp_data$developed %>% table()
```

Using this cutoff, we have 116 developed countries, and 122 developing countries.

```{r}
gdp_data %>% filter(developed==1) %>% head()
gdp_data %>% filter(developed==0) %>% head()
```

We then calculate GDP growth from year to year and finally an average GDP growth.

```{r}
gdp_data %>% mutate(growth10_11=(X2011-X2010)/X2010*100,
                    growth11_12=(X2012-X2011)/X2011*100,
                    growth12_13=(X2013-X2012)/X2012*100,
                    growth13_14=(X2014-X2013)/X2013*100,
                    growth14_15=(X2015-X2014)/X2014*100,
                    growth15_16=(X2016-X2015)/X2015*100,
                    avg_growth=(growth10_11+growth11_12+growth12_13+growth13_14+growth14_15+growth15_16)/6) -> gdp_data
```

```{r}
final_data %>% left_join(gdp_data, by=c("source.country"="Country.Name")) %>% filter(!is.na(Country.Code)) %>% left_join(gdp_data, by=c("destination.country"="Country.Name")) %>% filter(!is.na(Country.Code.y)) -> final_data; nrow(final_data)
```

42495->33559, where X2010.x...developed.x relate to source country GDP data, and X2010.y...developed.y relate to destination country GDP data.

### Addition of Population Data for Normalization:

```{r}
pop <- read.csv("Data/CountryPopulations2018.csv")
pop <- pop %>% mutate(Country = as.character(Country))
# 3 Country Names do not match:
pop <- pop %>% mutate(Country = ifelse(Country == "Ivory Coast", "Cote d'Ivoire",
                                       ifelse(Country == "Saint Kitts and Nevis", "St. Kitts and Nevis", ifelse(Country == "Saint Lucia", "St. Lucia", Country))))

hold <- left_join(final_data, pop, by = c("source.country" = "Country"))
colnames(hold)[46:47] <- c("source.population", "source.rank")
hold <- left_join(hold, pop, by = c("destination.country" = "Country"))
colnames(hold)[48:49] <- c("destination.population", "destination.rank")

final_data <- hold


```


There were 3 countries whose name was entered different between the population data and `final_data`. These were changed before the join.


```{r}
write.csv(final_data, "Data/final_data.csv")
```

