---
title: "Initial"
author: "Jocelyn, Daniel"
date: "3/1/2018"
output: html_document
---

```{r}
library(tidyverse)
library(stringr)
```

## Data

As of January 2012, the OpenFlights/Airline Route Mapper Route Database contains 67663 routes between 049 airports on 548 airlines spanning the globe. It was downloaded from https://www.kaggle.com/open-flights/flight-route-database. Our airport codes data was downloaded from http://www.airportcodes.org/#international. 

```{r}
data <- read.csv("routes.csv"); nrow(data)
codes <- read.csv("codes.csv", col.names="data"); nrow(codes)

df_data <- as.data.frame(data)
```

## Variables

```{r}
data %>% names()
```

- Airline 2-letter (IATA) or 3-letter (ICAO) code of the airline.
- Airline ID Unique OpenFlights identifier for airline (see Airline).
- Source airport 3-letter (IATA) or 4-letter (ICAO) code of the source airport.
- Source airport ID Unique OpenFlights identifier for source airport (see Airport)
- Destination airport 3-letter (IATA) or 4-letter (ICAO) code of the destination airport.
- Destination airport ID Unique OpenFlights identifier for destination airport (see Airport)
- Codeshare "Y" if this flight is a codeshare (that is, not operated by Airline, but another carrier), empty otherwise.
- Stops Number of stops on this flight ("0" for direct)
- Equipment 3-letter codes for plane type(s) generally used on this flight, separated by spaces

```{r}
data %>% left_join(codes, by=c("source.airport"="ident")) -> temp
```


## Data Wrangling

```{r}
codes %>% extract(col=data, c("Airport", "Airport Code"), "(.+)\\((...)\\)") -> codes
sapply(codes$Airport, function(x){unlist(strsplit(x,'-',''))[[1]]}) -> codes$Airport
codes %>% extract(col=Airport, c("Airport", "Country"), "(.+), (.+) ") -> codes
na.omit(codes) -> codes
```

```{r}
colnames(codes) <- c("airport","country","airport_code")

hold <- left_join(data, codes, by = c("source.airport" = "airport_code")) %>%
  filter(!is.na(country))

colnames(hold)[10] <- "source.airport.name"
colnames(hold)[11] <- "source.country"

hold2 <- left_join(hold, codes, by = c("destination.apirport" = "airport_code")) %>% 
  filter(!is.na(country))

colnames(hold2)[12:13] <- c("destination.airport.name","destination.country")
```


```{r}
# data %>% head(15) -> temp
data %>% mutate(source.country = ifelse(data$source.airport %in% codes$`Airport Code`,
                                        codes[which(data$source.airport %in% codes$`Airport Code` ),]$Country,
                                        NA),
                destination.country = ifelse(data$destination.apirport %in% codes$`Airport Code`,
                                        codes[which(data$destination.apirport %in% codes$`Airport Code` ),]$Country,
                                        NA)) -> data


data[!is.na(data$source.country),] %>% nrow()

```
