---
title: "Data Read-In"
author: "Jocelyn, Daniel"
date: "3/1/2018"
output: pdf_document
---

```{r}
library(tidyverse)
require(mosaic)
```

## Data

As of January 2012, the OpenFlights/Airline Route Mapper Route Database contains 67663 routes between 49 airports on 548 airlines spanning the globe. It was downloaded from https://www.kaggle.com/open-flights/flight-route-database. Our airport codes data was downloaded from http://www.airportcodes.org/#international. GDP data was downloaded from https://data.worldbank.org/indicator/NY.GDP.PCAP.PP.CD. 

```{r}
data <- read.csv("data/routes.csv"); nrow(data)
codes <- read.csv("data/codes.csv", col.names="data"); nrow(codes)
gdp_data <- read_csv("data/gdp_data.csv") %>% data.frame(); nrow(codes)
```

### Routes

```{r}
data %>% names()
```

- Airline 2-letter (IATA) or 3-letter (ICAO) code of the airline.
- Airline ID Unique OpenFlights identifier for airline (see Airline).
- Source airport 3-letter (IATA) or 4-letter (ICAO) code of the source airport.
- Source airport ID Unique OpenFlights identifier for source airport (see Airport)
- Destination airport 3-letter (IATA) or 4-letter (ICAO) code of the destination airport.
- Destination airport ID Unique OpenFlights identifier for destination airport (see Airport)
- Codeshare "Y" if this flight is a codeshare (that is, not operated by Airline, but another carrier), empty otherwise.
- Stops Number of stops on this flight ("0" for direct)
- Equipment 3-letter codes for plane type(s) generally used on this flight, separated by spaces

```{r}
data %>% head(2)
```

### Codes

```{r}
codes %>% head(2)
```

### GDP

```{r}
gdp_data %>% names()
```

- Country.Name
- Country.Code
- X2010:X2016

```{r}
gdp_data %>% head(2)
```


## Data Wrangling

```{r}
codes %>% extract(col=data, c("Airport", "Airport_Code"), "(.+)\\((...)\\)") -> codes
sapply(codes$Airport, function(x){unlist(strsplit(x,'-',''))[[1]]}) -> codes$Airport
codes %>% extract(col=Airport, c("Airport", "Country"), "(.+), (.+) ") -> codes
na.omit(codes) -> codes; nrow(codes)
```

Lose some codes 2508->2477

```{r}
temp <- left_join(data, codes, by = c("source.airport" = "Airport_Code")) %>%
  filter(!is.na(Country))

colnames(temp)[10:11] <- c("source.airport.name","source.country")

final_data <- left_join(temp, codes, by = c("destination.apirport" = "Airport_Code")) %>% 
  filter(!is.na(Country))

colnames(final_data)[12:13] <- c("destination.airport.name","destination.country")

nrow(final_data)
```

67663 -> 42495

```{r}
sapply(gdp_data[,3:9], function(x){which(is.na(x)) %>% length()})
is.na(gdp_data$X2010) -> missing
gdp_data[!missing,] -> gdp_data; nrow(gdp_data)
```

We have the most data for GDP  (PPP) figures for year 2010, so we'll proceed with this year. We now have n=238.

Using the inflation calculator provided by the BLS, https://data.bls.gov/cgi-bin/cpicalc.pl. \$12000 today would have been worth \$10490.48 in 2010.

```{r}
gdp_data$developed <- ifelse(gdp_data$X2010>10490.48, 1, 0);gdp_data$developed %>% table()
```

Using this cutoff, we have 116 developed countries, and 122 developing countries.

```{r}
gdp_data %>% filter(developed==1) %>% head()
gdp_data %>% filter(developed==0) %>% head()
```

We then calculate GDP growth from year to year and finally an average GDP growth.

```{r}
gdp_data %>% mutate(growth10_11=(X2011-X2010)/X2010*100,
                    growth11_12=(X2012-X2011)/X2011*100,
                    growth12_13=(X2013-X2012)/X2012*100,
                    growth13_14=(X2014-X2013)/X2013*100,
                    growth14_15=(X2015-X2014)/X2014*100,
                    growth15_16=(X2016-X2015)/X2015*100,
                    avg_growth=(growth10_11+growth11_12+growth12_13+growth13_14+growth14_15+growth15_16)/6) -> gdp_data
```

```{r}
gdp_data %>% ggplot(aes(avg_growth)) +  geom_histogram()
```


```{r}
final_data %>% left_join(gdp_data, by=c("source.country"="Country.Name")) %>% filter(!is.na(Country.Code)) %>% left_join(gdp_data, by=c("destination.country"="Country.Name")) %>% filter(!is.na(Country.Code.y)) -> final_data; nrow(final_data)
```

42495->33559, where X2010.x...developed.x relate to source country GDP data, and X2010.y...developed.y relate to destination country GDP data.




```{r}
write.csv(final_data, "data/final_data.csv")
```

